---
output:
  rmarkdown::github_document: default
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "fig/README-"
)
```

# rlena
[![Project Status: Concept - Minimal or no implementation has been done yet.](http://www.repostatus.org/badges/0.1.0/concept.svg)](http://www.repostatus.org/#concept)

An R package for extracting data from LENA .its files.

## Proof of Concept

There are just a few functions at the moment: 

- `read_its_file()` loads the ITS file in R as an XML document.
- `gather_ava_info()` extracts the AVA (Automatic Vocalization Assessment) scores into a dataframe.
- `gather_child_info()` extracts the child information.
- `gather_conversations()` collects the attributes of every `<Conversation>` node into a dataframe.

```{r, warning = FALSE}
library(rlena)
library(dplyr, warn.conflicts = FALSE)

lena_log <- read_its_file("tests/testthat/data/file-for-testing.its") 
```

Examples of simple information extraction.

```{r}
gather_ava_info(lena_log)
gather_child_info(lena_log)
```

We can extract a dataframe of conversations and play with the data.

```{r conversation-demo}
conversations <- gather_conversations(lena_log)
conversations

# Create long data-frame of word counts
library("tidyr")
word_counts <- conversations %>% 
  select(num, Time = startTime, 
         Female = femaleAdultWordCnt, Male = maleAdultWordCnt) %>% 
  gather(Speaker, Count, -num, -Time) %>% 
  filter(Count != 0)

library("ggplot2")
ggplot(word_counts) + 
  aes(x = Time, y = Count, color = Speaker) + 
  geom_point() + 
  ylab("Adult Word Count Per Conversation")
```



