---
output:
  rmarkdown::github_document: default
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "fig/README-"
)
```

# rlena
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)


## Overview
The `rlena` package makes it easy to work with LENA `.its` files in R.

The *Language Environment ANalysis (LENA)* system produces automatic annotations for audio recordings. Its annotations can be exported as `.its` files, that contain an `xml` structure. The `rlena` package helps with importing `.its` files into R and with creating and producing tidy data frames for further analysis.

The `read_its_file()` fuction takes the path to an .its file and returns an XML object.

All other functions work with this object to extract information:

  - `gather_ava_info()` returns the AVA (Automatic Vocalization Assessment).
  - `gather_child_info()` returns child information (e.g. birth data, age, gender).
  - `gather_conversations()` returns the content of all `<Conversation>` 
  nodes.
  
All three functions return tidy data frames that can be manipulated with the usual tools from the tidyverse.

***Warning:** Please be aware that this package is early work in progress and subject to potentially  code-breaking change.*

## Installation

```{r eval=FALSE}
# rlena is not (yet) available on CRAN. 
# You can install the developmental version from GitHub:

# install.packages("devtools")
devtools::install_github("HomeBankCode/rlena")
```


## Usage

Load an `.its` file

```{r, warning = FALSE}
library(rlena)
library(dplyr, warn.conflicts = FALSE)

# Loading an example .its file
its_file <- system.file('extdata', 'example.its', package = 'rlena')
its <- read_its_file(its_file) 
```

Extract child info and results of the Automatic Vocalization Assessment (an index of the child's language development).

```{r}
gather_child_info(its)
gather_ava_info(its)
```

Extract all conversations. Each row of the returned data frame corresponds to one converstion node in the `.its` file. The columns contain the node's attributes, such as the number of adult words and child vocalizations.

```{r}
conversations <- gather_conversations(its)
conversations
```

Plot male and female adult word counts.

```{r conversation-demo, fig.width = 6, fig.height = 3}
library("tidyr")

# Create long data-frame of word counts
word_counts <- conversations %>% 
  select(conversation_nr = num,
         time = startTime,
         female = femaleAdultWordCnt, 
         male = maleAdultWordCnt) %>% 
  gather(key = speaker, value = count, female, male) %>% 
  filter(count != 0)

# Add acumulated word count
word_counts <- word_counts %>%
  group_by(speaker) %>% 
  arrange(conversation_nr) %>%
  mutate(count_acc = cumsum(count))

library("ggplot2")
word_counts %>%
  ggplot(aes(time, count, color = speaker)) + 
    geom_point() + 
    labs(title = "Adult Word Count per Conversation",
         x = "Time (sec since start of recording)",
         y = "Words")
  
  ggplot(word_counts) + 
    aes(time, y = count_acc, color = speaker) + 
    geom_line() + 
    labs(title = "Accumulated Adult Word Count throughout Recording",
         x = "Time (sec since start of recording)",
         y = "Words (Accumulated)")
```
