---
title: "Converting XML to YAML"
author: "Tristan Mahr"
date: "February 4, 2016"
output: 
  md_document: 
    variant: markdown_github
---

Data from a LENA analysis is dumped inside of the `.its` file (ITS: Interpreted 
Time Segment). This file is a gigantic xml file, meaning that it stores data in 
a hierarchical tree. This is an inconvenient format, so let's convert the data 
into the comparatively simpler YAML format, where everything is a list of
key-value pairs.


```{r}
knitr::opts_knit$set(root.dir = "../")
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

Here's what the ExportData node in the its file looks like in XML:

```xml
<ExportData ProductRegistration="555-000000">
    <Child id="X555" EnrollDate="2001-01-01" ChildKey="2130A13191454FB7" DOB="1997-07-01" Gender="X" />
    <Group id="" name="" />
</ExportData>
```

We can use the XML package to parse the its-file into an R list.

```{r}
library("XML")
its_lines <- readr::read_lines("./tests/testthat/data/file-for-testing.its")
its_parse <- xmlTreeParse(its_lines) 
its_list <- xmlToList(its_parse)
str(its_list$ExportData)
```

And then use the yaml package to get the yaml format:

```{r}
library("yaml")
cat(as.yaml(its_list$ExportData))
```

But we've lost the names of the XML attributes in the sub-lists and in the 
`.attrs` list. We should convert attributes into lists so every value as a name.
I wrote some functions to recursively update the list.

```{r}
library("purrr")
source("./R/recurse.R")

# First wrap it up in a list. Next recursively convert items into lists and
# promote the values store in .attrs
export_data <- list(ExportData = its_list$ExportData)
yaml_export <- export_data  %>% 
  descend(promote_attributes) %>% 
  as.yaml
```

Finally, we have all the data for this node in a sensible format.

```{r}
yaml_export %>% cat
```

Now we can generalize this approach to other parts of the document.

```{r}
ava <- list(AVA = its_list$ProcessingUnit$AVA)
ava %>% 
  descend(promote_attributes) %>% 
  as.yaml %>% 
  cat
```

```xml
<AVA rawScore="-0.285" standardScore="95.727" estimatedMLU="3.286" estimatedDevAge="P43M" vocalizationCnt="ORH" vocalizationLen="ORH" MLV="ORH" />
```

```{r}
hour <- list(Bar = its_list$ProcessingUnit$Bar$BarSummary)
hour %>% 
  descend(promote_attributes) %>% 
  as.yaml %>% 
  cat
```

```xml
<BarSummary average_dB="-24.10" peak_dB="-7.73" adultWordCnt="791.00" turnTaking="48" childUttCnt="168" childUttLen="P147.02S" TVF="P197.67S" FAN="P171.01S" OLN="P689.08S" SIL="P10.53S" NOF="P208.94S" CXF="P37.92S" OLF="P422.14S" CHF="P3.00S" MAF="P5.25S" TVN="P533.50S" NON="P106.48S" CXN="P130.32S" CHN="P163.08S" MAN="P19.28S" FAF="P18.80S" totalRecordingTime="P2717.00S" leftBoundaryClockTime="2000-12-28T13:14:43Z" rightBoundaryClockTime="2000-12-28T14:00:00Z" pauseTime="P0.00S" />
```
